\emph{MixFishSim}: highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-07-18 write_up/mixfishsim_ecolmod.tex
Clean up abstract
mixfish corrections

\replaced[id = PD]{If fisheries are to avoid unwanted catch}{To implement
	effective spatial measures to reduce discards} a good understanding of
spatiotemporal fishery dynamics is required. However, traditional scientific
advice is limited by a lack of highly resolved knowledge of population
distributions, population movement and how fishers interact with different fish
populations. This reflects the fact that data on fish location at high temporal
and spatial resolutions is expensive and difficult to collect. Proxies inferred
from either scientific surveys or commercial catch data are often used to model
distributions, usually with sparse data at limited spatial and temporal
resolution. \\

To understand how data resolution impacts inference on mixed fisheries
interactions, we develop a highly resolved spatiotemporal simulation model
incorporating: i) delay-difference population dynamics, ii) population movement
using Gaussian Random Fields to simulate patchy, heterogeneously distributed
populations, and iii) fishery dynamics for multiple fleet characteristics based
on species targeting under an explore-exploit strategy via a mix of correlated
random walk movement (for exploration) and learned behaviour (for exploitation)
phases of the fisheries.  \\

Our framework allows users to explore the assumptions in modelling observational
data and evaluate the underlying dynamics of such approaches at a fine spatial
and temporal scale. \\

We simulate 50 years of fishing and use the results from the fisheries catch to
draw inference on the underlying population structures. We compare this
inference to a simulated fixed-site sampling design commonly used for fisheries
monitoring purposes and the true underlying population structures. We i) use
the results to establish the potential and limitations of fishery-dependent
data in providing a robust picture of spatiotemporal distributions; and ii)
simulate an area closure based on areas defined from the different data sources
at a range of temporal and spatial resolutions and assess their effectiveness
on reducing catches of a fish population. \\

We conclude from our simulations that commercial data, while containing bias,
provides a useful tool for managing catches in mixed fisheries if applied at
the correct spatiotemporal scale. \\

[333 words]

@@

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time, with varying knowledge of species distributions
and using species non-selective fishing gear. \added[id=PD]{In doing so}
fisheries \deleted[id=PD]{that}catch an assemblage of species
\added[id=PD]{and}\deleted{, known as mixed fisheries} \replaced[id=JJ]{may}{,
	when managed by single-species quotas can end up}
discard\deleted[id=JJ]{ing} over-quota catch \added[id=JJ]{when managed by
	single species quotas and fishers exhaust one or more quota. This may}
lead to overexploitation of fish populations \added[id=JJ]{\citep{Ulrich2011a,
		Batsleer2015}}.  \added[id=JJ]{Discarding of fish in
	excess of quota limits the ability to maintain fishing mortality within
	sustainable limits \citep{Alverson1994, Crowder1998, Rijnsdorp2007}}
\deleted[id=PD]{; reducing discarding is crucial}and the ability to manage for
the biological and economic sustainability of fisheries\deleted[id=JJ]{ and
	implementation of an ecosystem approach to fisheries}. As such, there
is increasing interest in technical solutions such as gear and spatial closures
as measures to \replaced[id=PD]{reduce unwanted catch}{avoiding
	discard\added[id=JJ]{ing of fish}} \citep{Kennelly2002, Catchpole2008,
	Bellido2011, Cosgrove2019}.  \\

\replaced[id=PD]{Changes to spatial fishing patterns have}{Use of spatial
	management as a tools} been proposed as a method to reduce discards
\added[id=PD]{\citep{Holmes2011, Little2014, Dunn2014a}}.  \deleted[id=PD]{its}
Implementation of avoidance measures is, however, restricted by lack of
knowledge of fish and fishery spatiotemporal dynamics and understanding of the
scale at which processes become important for management. Understanding the
correct scale for spatial measures is crucial for implementation at a
resolution that ensures effective management \citep{Dunn2016} while minimising
economic impact.  For example, a scale that promotes species avoidance for
vulnerable or low quota species while allowing continuance of sustainable
fisheries for available quota species.\\

\replaced[id=PD]{Identifying}{Ensuring measures are implemented at} an
appropriate scale has been a challenge in the past that has led to ineffectual
measures with unintended consequences such as limited impact towards the
management objective or increased benthic impact on previously unexploited
areas (e.g. the cod closure in the North Sea
\citep{Rijnsdorp2001,Dinmore2003}). \replaced[id=PD]{M}{Since then m}ore
refined spatial information has \added[id=PD]{since} become available through
the combination of logbook and Vessel Monitoring System (VMS) data
\citep{Lee2010, Bastardie2010, Gerritsen2012, Mateo2016} and more real-time
spatial management has been possible \citep[e.g.][]{Holmes2011}.  Such
information is, however, derived from an inherently biased sampling programme,
targeted fishing, where fishers establish favoured fishing grounds through an
explore-exploit strategy \citep{Bailey2018} where they search for areas with
high catches and then use experience to return to areas where they've
experienced high catch in the past.  \deleted[id=PD]{ Further, fishers
	generally only recorded landings (not catch) on a daily
	basis\todo{\added[id=JJ]{Its not clear what the problem is: landings
			collected on daily basis or landings recorded rather
			than catch}}.  This leads to questions about the
	validity of inference that can be drawn from landings data assigned to
	VMS activity pings.}\todo{\added[id=JJ]{This comes as a surprise: I
		thought this was going to be about
		discards}\added[id=PD]{Agree, have removed this to avoid
		confusion}}\\

\replaced[id=PD]{We ask two fundamental questions regarding spatiotemporal
	inference derived from observational data:}{ and use the
	results}\deleted[id=PD]{ from the fishing model:}

\begin{enumerate}
	\item How does sampling-derived data reflects the underlying population
		structures? 	
	\item How does data aggregation and source impact on spatial fisheries
	management measures?
\end{enumerate}

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time\replaced[id=PD]{. Being}{ and
	are} known \added[id=PD]{directly} rather than inferred from sampling
or commercial catch\added[id=PD]{, we can use the population model to validate
	how inference from fisheries-dependent and fisheries independent
	sampling relates to the real population structure in a way we could not
	with real data}. We ii) compare, at different spatial and temporal
aggregations, the 'real population' distributions to samples from
fisheries-dependent and fisheries independent catches to test if these are a
true reflection of the relative density of the populations. We then iii)
simulate a fishery closure to protect a species based on different spatial and
temporal data aggregations. We use these evaluations to draw inference on the
utility of commercial data in supporting management decisions.
\todo{\added[id=JJ]{If the paper has two goals this should be clear from the
		start, but may be better over 2 MSs}\added[id=PD]{I would like
		to keep both parts, but have made clearer in how its set out.
		The closure scenarios form validation of the data aggregation,
		rather than effectiveness of the closures themselves - so its a
		continuation of the same question in my eyes}}\deleted[id=PD]{.
	Further, we extend our analysis to a range of spatial and temporal
	scales to assess the impact of these processes on the success of the
	management measure.}\\

[We find..]

@@

The basic population level processes are simulated using a modified two-stage
Deriso-Schnute delay difference model \citep{Deriso1980, Schnute1985,
	Dichmont2003} occurring at a daily time-step. \added[id=PD]{A daily
	time-step was chosen to discretise continuous population processes on a
	biologically relevant and computationally tractable timescale.}
\replaced[id=PD]{Under the population dynamics module}{Here,} population
biomass growth and depletion for pre-recruits and\deleted[id=PD]{fish}
recruited \added[id=PD]{fish}\deleted[id=PD]{to the fishery} are modelled
separately as a function of previous recruited biomass, intrinsic population
growth and recruitment functionally linked to the adult population size.
Biomass for each cell $c$ is incremented each day $d$ as follows (the full
parameter list is detailed in Table \ref{tab: pop_variable}): \begin{equation}
	\begin{split}
	B_{c,d+1} = &\\
	& (1 + \rho) B_{c,d} \cdot e^{-Z_{c,d}} - \rho \cdot e^{-Z_{c,d}} \hspace{2.9cm}
	\times \\  
	& (B_{c,d-1} \cdot e^{-Z_{c,d-1}} + Wt_{R-1} \cdot \alpha_{d-1} \cdot
	R_{\tilde{y}(c,y,d-1)})
	\hspace{0.4cm} + \\
	& Wt_{R} \cdot \alpha_{d} \cdot R_{\tilde{y}(c,y,d)} 
	\end{split}
\end{equation}
where $\rho$ is Brody's coefficient, shown to be equal to $e^{-K}$ when $K$ is
the growth rate from a von Bertalanffy logistic growth model
\citep{Schnute1985}. $Wt_{R-1}$ is the average weight of fish prior to
recruitment, while $Wt_{R}$ is the average recruited weight. $\alpha_{d}$
represents the proportion of fish recruited during that day for the year, while
$R_{c,\tilde{y}}$ is the annual recruits in cell $c$ for year $y$. \\

Mortality $Z_{c,d}$ can be decomposed to natural mortality, $M_{c,d}$, and
fishing mortality, $F_{c,d}$, where both $M_{c,d}$ and $F_{c,d}$ are
instantaneous rates with $M_{c,d}$ fixed and $F_{c,d}$ calculated by solving
the Baranov catch equation \citep{Hilborn1992b} for $F_{c,d}$:
\begin{equation}
C_{c,d} = \frac{F_{c,d}}{F_{c,d} + M_{c,d}} * (1 - e^{-(F_{c,d} + M_{c,d})}) *
B_{c,d}
\end{equation}
where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. Here, catch and fishing
mortality are the sum of those across all fleets and vessels, where $F_{fl, v,
	c, d, p} = E_{fl, v, c, d} \cdot Q_{fl, p} \cdot D_{c, d, p}$ with
$fl$, $v$ and $p$ the fleet, vessel and population respectively and $E$ and $Q$
fishing effort and catchability of the gear, and $D$ is the density of the
population at the location fished. \todo{\added[id=CM]{[link $F$ to effort and
		catchability - as I think we have F as an emergent property of
		the fleets rather than something we solve for (I could be wrong
		though!) - catch for a vessel is a product of catchability and
		biomass, i.e. C = qB, but this catch is summed to solve for F.
		So its both really]}}\\

@@

Recruitment is modelled through a function relating the adult biomass to
recruits at time of recruitment. In \emph{MixFishSim}, it can be modelled
either either as a stochastic Beverton-Holt stock-recruit form
\citep{Beverton1957}: 
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha * S_{c,d})}{(\beta + S_{c,d})} \\
	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\sigma^2)]
	\end{split}
\end{equation}
Where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum stock size, $S$ current
stock size and $\sigma^2$ the variability in the recruitment due to stochastic
processes, or a stochastic Ricker form \citep{Ricker1954}:
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} * e^{(\alpha - \beta * B_{c,d})} \\	
   	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the density
dependent reduction in productivity as the SSB increases. In our example
application the Beverton-Holt form of stock recruit relationship was used for
all populations though either functional form can be chosen.

@@

To simulate \deleted[id=JJ]{how} fish population\deleted[id=JJ]{s might be}
distribut\replaced[id=JJ]{ion}{ed} in space and time\deleted[id=JJ]{, we
	employed} a Gaussian spatial process \added[id=JJ]{was employed} to
model habitat suitability for each of the populations on a 2d grid. \\

@@

\replaced[id=PD]{T}{In the simulation model, t}he habitat for each of the
populations \replaced[id=PD]{was}{is} generated \replaced[id=PD]{with}{through}
the \textit{RFSimulate} function of the \textit{RandomFields} R package
\citep{Schlater2015}, that simulates a Gaussian Random Field process given a
user defined error model and correlation structure. We define a stationary
habitat field and combine with a temporally dynamic thermal tolerance field to
imitate two key drivers of population dynamics. Each population
\replaced[id=PD]{was}{is} initialised at a single location, and subsequently
move\replaced[id=PD]{ed}{s} according to a probabilistic
distrib\added[id=PD]{u}tion based on habitat suitability (represented by the
normalised values from the GRFs), temperature and distance
from current cell\replaced[id=PD]{:}{.} 
\begin{equation}
	Pr(J | I) = \frac{e^{-\lambda * d_{IJ}} \cdot
		(Hab_{J,p}^2 \cdot Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}e^{-\lambda * d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{IJ}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{J,p}^2$ is the squared index of
habitat suitability for cell $J$ and population $p$, with $Tol_{J,p,wk}$ the
temperature tolerance for cell $J$ by population $p$ in week $wk$ (see
below).\\

During pre-defined weeks of the year the habitat quality is modified with
\added[id=PD]{user-defined} spawning habitat locations\deleted[id=PD]{s},
\replaced[id=PD]{resulting in}{meaning} each population
\replaced[id=PD]{having}{has} concentrated areas where spawning takes place. In
the simulations the populations move\deleted[id=PD]{s} towards
\replaced[id=PD]{these cells}{this} in the weeks prior to spawning, resulting
in directional movement towards the spawning grounds.\todo{\added[id=JM]{What
		does it mean concisely? Areas are assigned?} \added[id=PD]{Yes,
		the areas are pre-defined - I have amended to reflect and tried
		to clarify}} \\

@@

The fleet dynamics can be broadly categorised into three components; fleet
targeting - that determines\replaced[id=PD]{d}{s} the fleet catch efficiency
and preference towards a particular species; trip-level decisions, that 
determines\added[id=PD]{d} the initial location to be fished at the beginning of
a trip; and within-trip decisions, determining movement from one fishing spot
to another within a trip. Together, these elements implement an explore-exploit
type strategy for individual vessels to maximise their catch from an unknown
resource distribution \cite{Bailey2018}. The decision to use an individual
based model for fishing vessels was taken because fishers are heterogeneous in
their location choice behaviour due to different objectives, risk preference
and targeting preference \citep{VanPutten2012a}. Therefore in the simulations
fleet dynamics are the productive of individual experiences rather than
pre-defined group dynamics.

@@

Where several fishing vessels exploit the same fish population competition is
known to play an important role in local distribution of fishing effort
\citep{Gillis1998}. If several vessels are fishing on the same patch of fish,
local depletion and interference \added[id=JJ]{competition} will affect fishing
location choice of the fleet as a whole \citep{Rijnsdorp2000, Poos2007a}. In
order to account for this behaviour, the fishing sub-model operates spatially
on a daily time-step so that for future days the biomass available to the
fishery is reduced in the areas fished. The cumulative effect is to make
heavily fished areas less attractive as a future fishing location choice as
reduced catch rates will be experienced.

@@

\subsection{Software}

@@

We parametrised the simulation model for four populations with different
demographics; growth rates, natural mortality and recruitment functions (Table
\ref{tab:1}). Habitat preference (Figure \ref{S-fig:1}) and temperature
tolerances (Figures \ref{S-fig:3}, \ref{S-fig:4}) were unique to each
population resulting in differently weekly distribution patterns (Figures
\ref{S-fig:5}-\ref{S-fig:7}). In addition, each of the populations has two
defined spawning areas that result in the populations moving towards these
areas in pre-defined weeks (Figure \ref{S-fig:2}) with population-specific
movement rates (Table \ref{tab:1}). The individual habitat preferences and
thermal tolerances result in different spatial habitat use for each population
(Figure \ref{S-fig:9}) and consequently different seasonal exploitation
patterns (Fishing mortality in Figure \ref{S-fig:10}).

@@

The fleets were parametrised to reflect five different characteristic fisheries
with unique exploitation dynamics (Table \ref{tab:2}).  \replaced[id=PD]{By
	setting different catchability parameters ($Q_{fl, p}$) we create
	different targeting preferences between the fleets and hence spatial
	dynamics}{This ensures that different fleets have different spatial
	dynamics, preferentially targeted different fish populations}.  The
stochasticity in the random walk process ensures that within a fleet different
vessels have slightly different spatial distributions based on individual
experience. The step function was parametrised dynamically within the
simulations as the maximum revenue obtainable was not known beforehand. This
was implemented so that vessels take smaller steps when fishing at a location
that yields landings value in the top 90th percentile of the value experienced
in that year so far (as defined per fleet in Table \ref{tab:2}). \\

With increasing probability throughout the simulation, fishing locations were
chosen based on experience of profitable catches built up in the same month
from previous years and from the previous trip. 'Profitable' in this context
was defined as the locations where the top 70 \% of expected profit would be
found given previous trips revenue and cost of movement to the new fishing
location.  This probability was based on a logistic sigmoid function with a
lower asymptote of 0 and upper asymptote of 0.95, and a growth rate that 
ensures the upper asymptote (where decisions are mainly based on past
knowledge) is reached approximately halfway through the simulation.  \\

@@

%%%% figures
\begin{figure}[!ht]
	\includegraphics[width=\linewidth]{../analysis/Data_Aggregation_space_Rev}
	\caption{Data aggregation at different spatial resolutions over a
		ten year period}
	\label{fig:1}
\end{figure}

>>>>

\replaced[id = PD]{If fisheries are to avoid unwanted catch,}{To implement
	effective spatial measures to reduce discards} a good understanding of
spatiotemporal fishery dynamics is required. However, traditional scientific
advice is limited by a lack of highly resolved knowledge of population
distributions, population movement and how fishers interact with different fish
populations. This reflects the fact that data on fish location at high temporal
and spatial resolutions is expensive and difficult to collect. Proxies inferred
from either scientific surveys or commercial catch data are often used to model
distributions, usually with sparse data at limited spatial and temporal
resolution. \\

To understand how data resolution impacts inference on mixed fisheries
interactions, we developed a highly resolved spatiotemporal simulation model
incorporating: i) delay-difference population dynamics, ii) population movement
using Gaussian Random Fields to simulate patchy, heterogeneously distributed
populations, and iii) fishery dynamics for multiple fleet characteristics based
on species targeting under an explore-exploit strategy via a mix of correlated
random walk movement (for exploration) and learned behaviour (for exploitation)
phases of the fisheries.  \\

We simulated 50 years of fishing and used the results from the fisheries catch
to draw inference on the underlying community structures. We compared this
inference to a simulated fixed-site sampling design commonly used for fisheries
monitoring purposes and the true underlying community structure. We i) used
the results to establish the potential and limitations of fishery-dependent
data in providing a robust picture of spatiotemporal distributions; and ii)
simulated an area closure based on areas defined from the different data sources
at a range of temporal and spatial resolutions to assess their effectiveness
on reducing catches of a fish population. \\

Our framework allows users to explore the assumptions in modelling
observational data and evaluate the underlying dynamics of such approaches at a
fine spatial and temporal scale. We conclude from our simulations that
commercial data, while containing bias, provide a useful tool for managing
catches in mixed fisheries if applied at the correct spatiotemporal scale. \\

@@

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time, with varying knowledge of species
distributions\deleted[id=JJ]{ and using species non-selective fishing gear}.
\added[id=PD]{In doing so} fisheries \deleted[id=PD]{that}catch an assemblage
of species \added[id=PD]{and}\deleted{, known as mixed fisheries}
\replaced[id=JJ]{may}{, when managed by single-species quotas can end up}
discard\deleted[id=JJ]{ing} over-quota catch \added[id=JJ]{when managed by
	single species quotas and fishers exhaust one or more quota. This may}
lead to overexploitation of fish populations \added[id=JJ]{\citep{Ulrich2011a,
		Batsleer2015}}.  \added[id=JJ]{Discarding of fish in
	excess of quota limits the ability to maintain fishing mortality within
	sustainable limits \citep{Alverson1994, Crowder1998, Rijnsdorp2007}}
\deleted[id=PD]{; reducing discarding is crucial}and the ability to manage for
the biological and economic sustainability of fisheries\deleted[id=JJ]{ and
	implementation of an ecosystem approach to fisheries}. As such, there
is increasing interest in technical solutions such as gear and spatial closures
as measures to \replaced[id=PD]{reduce unwanted catch}{avoiding
	discard\added[id=JJ]{ing of fish}} \citep{Kennelly2002, Catchpole2008,
	Bellido2011, Cosgrove2019}.  \\

\replaced[id=PD]{Adaptive spatial management strategies have}{Use of spatial
	management as a tools} been proposed as a way of reducing discards
\added[id=PD]{\citep{Holmes2011, Little2014, Dunn2014a}}.  \deleted[id=PD]{its}
Implementation of avoidance measures is, however, restricted by lack of
knowledge of fish and fishery spatiotemporal dynamics and understanding of the
scale at which processes become important for management. Understanding the
correct scale for spatial measures is crucial for implementing solutions at a
resolution that ensures effective management \citep{Dunn2016} while minimising
economic impact.  For example, the problem can be to identify a scale that
promotes species avoidance for vulnerable or low quota species while allowing
continuance of sustainable fisheries for available quota species.\\

\replaced[id=PD]{Identifying}{Ensuring measures are implemented at} appropriate
scales for spatial fisheries management has been a challenge in the past that
has led to ineffectual measures with unintended consequences. These unintended
consequences limited impacts towards management objectives, or increased
benthic impact on previously unexploited areas (e.g.  the cod closure in the
North Sea \citep{Rijnsdorp2001,Dinmore2003}). SENTENCE HERE ATTRIBUTING THE
CHALLENGE OF IDENTIFYING APPROPRIATE SCALES TO COURSE SPATIAL INFORMATION.
\replaced[id=PD]{M}{Since then
	m}ore refined spatial information has \added[id=PD]{since} become
available through the combination of logbook and Vessel Monitoring System (VMS)
data \citep{Lee2010, Bastardie2010, Gerritsen2012, Mateo2016} and more
real-time spatial management has been possible \citep[e.g.][]{Holmes2011}.
Such
information is, however, derived from an inherently biased sampling programme,
targeted fishing, where fishers establish favoured fishing grounds through an
explore-exploit strategy \citep{Bailey2018} where they search for areas with
high catches and then use experience to return to areas where they've
experienced high catch in the past.  \deleted[id=PD]{ Further, fishers
	generally only recorded landings (not catch) on a daily
	basis\todo{\added[id=JJ]{Its not clear what the problem is: landings
			collected on daily basis or landings recorded rather
			than catch}}.  This leads to questions about the
	validity of inference that can be drawn from landings data assigned to
	VMS activity pings.}\todo{\added[id=JJ]{This comes as a surprise: I
		thought this was going to be about
		discards}\added[id=PD]{Agree, have removed this to avoid
		confusion}}\\

\replaced[id=PD]{In order to understand the effect of spatiotemporal
	aggregation of data we ask two fundamental questions regarding
	inference derived from observational data:}{ and use the
	results}\deleted[id=PD]{ from the fishing model:}

\begin{enumerate}
	\item How does sampling-derived data reflects the underlying community 
		structure? 	
	\item How does data aggregation and source impact on spatial fisheries
	management measures?
\end{enumerate}

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time\replaced[id=PD]{. Being}{ and
	are} known \added[id=PD]{directly} rather than inferred from sampling
or commercial catch\added[id=PD]{, we can use the population model to validate
	how inference from fisheries-dependent and fisheries independent
	sampling relates to the real community structure in a way we could not
	with real data}. We ii) compare, at different spatial and temporal
aggregations, the 'real population' distributions to samples from
fisheries-dependent and fisheries independent catches to test if these are a
true reflection of the relative density of the populations. We then iii)
simulate a fishery closure to protect a species based on different spatial and
temporal data aggregations. We use these evaluations to draw inference on the
utility of commercial data in supporting management decisions.
\todo{\added[id=JJ]{If the paper has two goals this should be clear from the
		start, but may be better over 2 MSs}\added[id=PD]{I would like
		to keep both parts, but have made clearer in how its set out.
		The closure scenarios form validation of the data aggregation,
		rather than effectiveness of the closures themselves - so its a
		continuation of the same question in my eyes}}\deleted[id=PD]{.
	Further, we extend our analysis to a range of spatial and temporal
	scales to assess the impact of these processes on the success of the
	management measure.}\\

@@

The basic population level processes were simulated using a modified two-stage
Deriso-Schnute delay difference model \citep{Deriso1980, Schnute1985,
	Dichmont2003} occurring at a daily time-step. \added[id=PD]{A daily
	time-step was chosen to discretise continuous population processes on a
	biologically relevant and computationally tractable timescale.}
\deleted[id=PD]{Here,}Population
biomass growth and depletion for pre-recruits and\deleted[id=PD]{fish}
recruited \added[id=PD]{fish}\deleted[id=PD]{to the fishery} were modelled
separately as a function of previous recruited biomass, intrinsic population
growth and recruitment functionally linked to the adult population size.
Biomass for each cell $c$ was incremented each day $d$ as follows (the full
parameter list is detailed in Table \ref{tab: pop_variable}): 
\begin{equation}
	\begin{split}
	B_{c,d+1} = &\\
	& (1 + \rho) B_{c,d} \cdot e^{-Z_{c,d}} - \rho \cdot e^{-Z_{c,d}} \hspace{2.9cm}
	\times \\  
	& (B_{c,d-1} \cdot e^{-Z_{c,d-1}} + Wt_{R-1} \cdot \alpha_{d-1} \cdot
	R_{\tilde{y}(c,y,d-1)})
	\hspace{0.4cm} + \\
	& Wt_{R} \cdot \alpha_{d} \cdot R_{\tilde{y}(c,y,d)} 
	\end{split}
\end{equation}
where $\rho$ is Brody's coefficient, shown to be equal to $e^{-K}$ when $K$ is
the growth rate from a von Bertalanffy logistic growth model
\citep{Schnute1985}. $Wt_{R-1}$ is the average weight of fish prior to
recruitment, while $Wt_{R}$ is the average recruited weight. $\alpha_{d}$
represents the proportion of fish recruited during that day for the year, while
$R_{c,\tilde{y}}$ is the annual recruits in cell $c$ for year $y$. \\

Mortality $Z_{c,d}$ can be decomposed to natural mortality, $M_{c,d}$, and
fishing mortality, $F_{c,d}$, where both $M_{c,d}$ and $F_{c,d}$ are
instantaneous rates with $M_{c,d}$ fixed and $F_{c,d}$ calculated by solving
the Baranov catch equation \citep{Hilborn1992b} for $F_{c,d}$:
\begin{equation}
C_{c,d} = \frac{F_{c,d}}{F_{c,d} + M_{c,d}} * (1 - e^{-(F_{c,d} + M_{c,d})})
\cdot B_{c,d}
\end{equation}
where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. Here, catch and fishing
mortality are the sum of those across all fleets and vessels, where $F_{fl, v,
	c, d, p} = E_{fl, v, c, d} \cdot Q_{fl, p} \cdot D_{c, d, p}$ with
$fl$, $v$ and $p$ the fleet, vessel and population respectively and $E$ and $Q$
fishing effort and catchability of the gear, and $D$ is the density of the
population at the location fished. \todo{\added[id=CM]{[link $F$ to effort and
		catchability - as I think we have F as an emergent property of
		the fleets rather than something we solve for (I could be wrong
		though!) - catch for a vessel is a product of catchability and
		biomass, i.e. C = qB, but this catch is summed to solve for F.
		So its both really]}}\\

@@

Recruitment is modelled through a function relating the adult biomass to
recruits at time of recruitment. In \emph{MixFishSim}, it can be modelled
either either as a stochastic Beverton-Holt stock-recruit form
\citep{Beverton1957}: 
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha \cdot S_{c,d})}{(\beta + S_{c,d})} \\
	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\sigma^2)]
	\end{split}
\end{equation}
Where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum stock size, $S$ current
stock size and $\sigma^2$ the variability in the recruitment due to stochastic
processes, or a stochastic Ricker form \citep{Ricker1954}:
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} \cdot e^{(\alpha - \beta \cdot B_{c,d})} \\	
   	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the density
dependent reduction in productivity as the SSB increases. In our example
application the Beverton-Holt form of stock recruit relationship was used for
all populations though either functional form can be chosen.

@@

To simulate \deleted[id=JJ]{how} fish population\deleted[id=JJ]{s might be}
distribut\replaced[id=JJ]{ion}{ed} in space and time\deleted[id=JJ]{, we
	employed} a Gaussian spatial process \added[id=JJ]{was employed} to
model habitat suitability for each of the populations on a 2d grid. JM: MENTION
IN THE INTRODUCTION \\

@@

\replaced[id=PD]{T}{In the simulation model, t}he habitat for each of the
populations \replaced[id=PD]{was}{is} generated \replaced[id=PD]{with}{through}
the \textit{RFSimulate} function of the \textit{RandomFields} R package
\citep{Schlater2015}, that simulates a Gaussian Random Field process given a
user defined error model and correlation structure. We define a stationary
habitat field and combine with a temporally dynamic thermal tolerance field to
imitate two key drivers of population dynamics. Each population
\replaced[id=PD]{was}{is} initialised at a single location, and subsequently
move\replaced[id=PD]{d}{s} according to a probabilistic
distrib\added[id=PD]{u}tion based on habitat suitability (represented by the
normalised values from the GRFs), temperature and distance
from current cell\replaced[id=PD]{:}{.} 
\begin{equation}
	Pr(J | I) = \frac{e^{-\lambda * d_{IJ}} \cdot
		(Hab_{J,p}^2 \cdot Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}e^{-\lambda * d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{IJ}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{J,p}^2$ is the squared index of
habitat suitability for cell $J$ and population $p$, with $Tol_{J,p,wk}$ the
temperature tolerance for cell $J$ by population $p$ in week $wk$ (see
below).\\

During pre-defined weeks of the year the habitat suitability is modified with
\added[id=PD]{user-defined} spawning habitat locations\deleted[id=PD]{s},
\replaced[id=PD]{resulting in}{meaning} each population
\replaced[id=PD]{having}{has} concentrated areas where spawning takes place. In
the simulations the populations move\deleted[id=PD]{s} towards
\replaced[id=PD]{these cells}{this} in the weeks prior to spawning, resulting
in directional movement towards the spawning grounds.\todo{\added[id=JM]{What
		does it mean concisely? Areas are assigned?} \added[id=PD]{Yes,
		the areas are pre-defined - I have amended to reflect and tried
		to clarify}} \\ JM: WHAT ABOUT INDIVIDUAL INTERACTIONS:w

@@

The fleet dynamics can be broadly categorised into three components; fleet
targeting - that determined the fleet catch efficiency and preference towards a
particular species; trip-level decisions, that determined the initial location
to be fished at the beginning of a trip; and within-trip decisions, that
determined movement from one fishing spot to another within a trip. Together,
these elements implemented an explore-exploit type strategy for individual
vessels to maximise their catch from an unknown resource distribution
\cite{Bailey2018}.  The decision to use an individual based model for fishing
vessels was taken because fishers are heterogeneous in their location choice
behaviour due to different objectives, risk preference and targeting preference
\citep{VanPutten2012a}. Therefore in the simulations fleet dynamics are the
productive of individual experiences rather than pre-defined group dynamics.

@@

Where several fishing vessels exploit the same fish population competition is
known to play an important role in local distribution of fishing effort
\citep{Gillis1998}. If several vessels are fishing on the same patch of fish,
local depletion and interference \added[id=JJ]{competition} will affect fishing
location choice of the fleet as a whole \citep{Rijnsdorp2000, Poos2007a}. In
order to account for this behaviour, the fishing sub-model operates spatially
on a daily time-step so that for future days the biomass available to the
fishery is reduced in the areas fished. The cumulative effect is to make
heavily fished areas less attractive as a future fishing location choice as
reduced catch rates will be experienced. JM: INTERFERENCE COMPETITION COULD
ALSO BE REPRESENTED BY A LIMITATION FACTOR

@@

\subsection{Software: R-package development}

@@

We parametrised the simulation model for four populations with different
demographics; growth rates, natural mortality and recruitment functions (Table
\ref{tab:1}). Habitat preference (Figure \ref{S-fig:1}) and temperature
tolerances (Figures \ref{S-fig:3}, \ref{S-fig:4}) were defined to be unique to
each population resulting in differently weekly distribution patterns (Figures
\ref{S-fig:5}-\ref{S-fig:7}). In addition, each of the populations was assumed
to have two defined spawning areas that result in the populations moving
towards these areas in pre-defined weeks (Figure \ref{S-fig:2}) with
population-specific movement rates (Table \ref{tab:1}). In such a
configuration, the individual habitat preferences and thermal tolerances result
in different spatial habitat use for each population (Figure \ref{S-fig:9}) and
consequently different seasonal exploitation patterns (Fishing mortality in
Figure \ref{S-fig:10}).

@@

The fleets were parametrised to reflect five different characteristic fisheries
with unique exploitation dynamics (Table \ref{tab:2}).  \replaced[id=PD]{By
	setting different catchability parameters ($Q_{fl, p}$) we create
	different targeting preferences between the fleets and hence spatial
	dynamics}{This ensures that different fleets have different spatial
	dynamics, preferentially targeted different fish populations}.  The
random walk process implies that within a fleet different vessels have
different spatial distributions based on individual experience. The step
function was parametrised dynamically within the simulations as the maximum
revenue obtainable was not known beforehand. This was implemented so that
vessels take smaller steps when fishing at a location that yields landings
value in the top 90th percentile of the value experienced in that year so far
(as defined per fleet in Table \ref{tab:2}). \\

With increasing probability throughout the simulation, fishing locations were
chosen based on experience of profitable catches built up in the same month
from previous years and from the previous trip. 'Profitable' in this context
was defined as the locations where the top 70 \% of expected profit would be
found given revenue from previous trips and cost of movement to the new fishing
location. This probability was based on a logistic sigmoid function with a
lower asymptote of 0 and upper asymptote of 0.95, and a growth rate that
ensures the upper asymptote (where decisions are mainly based on past
knowledge) is reached approximately halfway through the simulation.  \\

@@

%%%%%%%%%%%%%%%%%
%%%% figures %%%%
%%%%%%%%%%%%%%%%%

\begin{figure}[!ht]
	\includegraphics[width=\linewidth]{../analysis/Data_Aggregation_space_Rev}
	\caption{Data aggregation at different spatial resolutions over a
		ten year period}
	\label{fig:1}
\end{figure}
